---
title: "GTD Graphs Notebook"
output: html_document
---



Read 2016 GTD database from Excel file download.
```{r options(warn=0)}
library(readxl)
gtd <- read_excel("~/Downloads/globalterrorismdb_0221dist.xlsx")


```


```{r options(warn=0)}
imagmiss <-
function(data,name="")
{

#Function to create a graph of the observations of the dataset
#leaving white lines where data is missing.
#The main idea is to use the original dataset to create 
#a temporary dataset containing 1 if a value is found or
#0 is the value is missing. The temporary data set is graphed by column
#changing color for each feature and leaving a blank horizontal line if
#value is missing.

#Uses the R function image from the base library.

#data:  the dataset
#name:the name of the dataset as desired in the graph title

  ncol=dim(data)[2]
  nrow=dim(data)[1]

  xaxis=colnames(data)
    xaxis = xaxis[-ncol]
    ticks = 1:(dim(data)[2] - 1)
    data = as.matrix(data)
    data = data[, -ncol]
    cat("Report on missing values for ",name,":\n")
    cat("\nNumber of missing values overall: ")
    cat(sum(is.na(data)))
    cat("\nPercent of missing values overall: ")
   cat((sum(is.na(data))/(dim(data)[1]*dim(data)[2]))*100)
   cat("\nFeatures with missing values (percent): \n")
   print(colSums(is.na(data))[colSums(is.na(data))!=0]/dim(data)[1]*100)
   #cat("\n",length(which(colSums(is.na(data))!=0)))
   cat("\nPercent of features with missing values: ")
   cat(length(which(colSums(is.na(data))!=0))/dim(data)[2]*100)
   cat("\nNumber of instances with missing values: ")
   cat(length(which(rowSums(is.na(data))!=0, arr.ind=TRUE)))
   cat("\nPercent of instances with missing values: ")
   cat(length(which(rowSums(is.na(data))!=0))/dim(data)[1]*100)
   cat("\n")
  data[which(data!="NA")]=1
  data[-which(data!="NA")]=0

  ncol1=ncol-1

  for(i in 1:ncol1)
   {
     data[data[,i]!=0,i]=i
   }

  x=1:ncol1
  y=1:nrow
data=apply(data,1,as.numeric)
  graph.title=paste("Distribution of missing values by variable for - ",name)
 
  image(x,y,data,col=c(0,topo.colors(100)),xlab="features",ylab="instances",axes=FALSE,main=(graph.title),cex.main=.7)
  axis(1,labels=xaxis,at=ticks)
}


```


```{r options(warn=0)}
imagmiss(gtd, name="Global Terrorism Database")

```




Create Table with number of terrorist incidents by country by year for top 20 countries.
```{r}
w<-table(gtd[,2]) #total world incidents by year
x<-table(gtd[,9]) # total number of incidents by country
x<-sort(x) # list of countries in ascending order by total number of incidents by country
top10<-rownames(tail(x,10)) # top 10 countries by number of terrorist incidents
t<-table(gtd[gtd$country_txt %in% top10,c(9,2)])
#t
#w

```

```{r}
t

```


Exponential Plot for number of incidents by year for Irak and the World.
```{r}
colors <- c("blue","red","green","black")
plot(colnames(t),t["Iraq",], log="y", type="l", col=colors[1], 
      ylim=c(1, 40000),
      xaxt="n",
      main = "Historical terrorist attacks in Iraq, Afghanistan and Pakistan",
      xlab="Year", ylab="Annual Terrorist Attacks")
lines(rownames(w),t["Afghanistan",], type="l", col=colors[2])#, pchl
lines(rownames(w),t["Pakistan",], type="l", col=colors[3])#, pchl
lines(rownames(w),w, type="l", col=colors[4])#, pchl

legend("bottomright", # places a legend at the appropriate place c(“Health”,”Defense”), # puts text in the legend
legend=c("World","Pakistan","Afghanistan","Iraq"),
  lty=c(1,1), # gives the legend appropriate symbols (lines)
  lwd=c(2.5,2.5),col=colors[4:1]) # gives the legend lines the correct color and width
grid(lty=2)
x_axis = rownames(w)
x_axis = x_axis[1:length(x_axis) %% 2 == 0]
axis(1,x_axis, las=2, cex.axis=.8)
```
Logarithmic Plot for terrorist attacks by year for top 10 countries and the world
```{r}
top<-rownames(tail(x,10)) # top 10 countries by number of terrorist incidents
t<-table(gtd[gtd$country_txt %in% top,c(9,2)])
colors <- rainbow(nrow(t))
countries = rownames(t)
plot(colnames(t),t["Iraq",], log="y", type="n", col="black",
     ylim=c(1, 40000), 
     main = "Terrorist attacks by year for top 10 countries\n(logarithmic graph)",
     xlab="Year", ylab="Number of Terrorist Attacks", xaxt="n")
lines(rownames(w), w, type="l", col="black")#, pchl
for(i in 1:length(countries)){
 lines(colnames(t),t[i,], type="l", col=colors[i])#, pchl
}


legend("topleft", # places a legend at the appropriate place c(“Health”,”Defense”), # puts text in the legend
  legend=c("World",countries),
  lty=c(1,1), # gives the legend appropriate symbols (lines)
  lwd=c(2.5,2.5),col=c("black",colors), cex=.7, ncol=4) # gives the legend lines the correct color and width
grid(lty=2)
axis(1,x_axis, las=2, cex.axis=.8)
```

Plot number of incidents by year for top 10 countries.
```{r}
top<-rownames(tail(x,10)) # top 10 countries by number of terrorist incidents
t<-table(gtd[gtd$country_txt %in% top,c(9,2)])
colors <- rainbow(nrow(t))
countries = rownames(t)
plot(colnames(t),t["Iraq",], type="n", col=colors[1],ylim=c(1, 4000),
     xaxt="n",
     main = "Terrorist attacks by year for top 10 countries",
     xlab="Year", ylab="Number of Terrorist Attacks")
for(i in 1:length(countries)){
  lines(colnames(t),t[i,], type="l", col=colors[i])#, pchl
}


legend("topleft", # places a legend at the appropriate place c(“Health”,”Defense”), # puts text in the legend
legend=countries,
lty=c(1,1), # gives the legend appropriate symbols (lines)
lwd=c(2.5,2.5),col=colors, cex=.8, ncol=1) # gives the legend lines the correct color and width
grid(lty=2)
axis(1,x_axis, las=2, cex.axis=.8)
```
Individual plots of number of incidents by year for top 10 countries.
```{r}
ncountries=10
top<-rownames(tail(x,ncountries)) # top 10 countries by number of terrorist incidents
t<-table(gtd[gtd$country_txt %in% top,c(9,2)])
t<-t[order(rank(rowSums(t))),]
colors <- rainbow(nrow(t))
countries = rownames(t)
#par(pch=22, col="red") # plotting symbol and color 
par(mfrow=c(ncountries/5,5)) # all plots on one page 
for(i in 1:length(countries)){
  plot(colnames(t),t[i,], type="l", col=colors[i],
       ylim=c(1, max(t[,])), 
       xaxt="n",
       xlab="Year", ylab="Number of Terrorist Attacks",main=countries[i])
grid(lty=2)
x_axis = rownames(w)
x_axis = x_axis[1:length(x_axis) %% 9 == 0]
axis(1,x_axis, las=2, cex.axis=.77)
}
```